### 2019-04-04 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Okay, okay, apparently « on the one hand » is (slightly) more frequent. https://dictionary.cambridge.org/fr/dictionnaire/anglais/on-the-one-hand-on-the-other-hand
Using \begin{framed}...\end{framed} for all algorithms.
Moved a lot of appendix in main text for Chapter 5. See #15

Multi-player Multi-Armed Bandits (MAB) have been extensively studied in the literature, motivated by applications to Cognitive Radio systems.
Driven by such applications as well, we motivate the introduction of several levels of feedback for multi-player MAB algorithms.
Existing works assume that \emph{sensing information} is available to the algorithm. Under this assumption, we improve the state-of-the-art lower bound for the regret of any decentralized algorithms of a certain class.
We introduce two algorithms, \RandTopM{} and \MCTopM{}, that are shown to empirically outperform existing algorithms. Moreover, we provide strong theoretical guarantees for these algorithms, including
a notion of asymptotic optimality in terms of the number of selections of bad arms.
We then introduce a promising heuristic, called \Selfish{}, that can operate
without sensing information, which is crucial for emerging applications to Internet of Things networks. We investigate the empirical performance of this algorithm
and provide some first theoretical elements for the understanding of its behavior.

@@

For the development of CR, a crucial step is to insert \emph{multiple} $M \geq 2$ smart devices in the \emph{same} background traffic.
With the presence of a central controller that can assign the devices to separate channels, this amounts to choosing at each time step \emph{several} arms of a MAB in order to maximize the global rewards, and can thus be viewed as an application of the multiple-play bandit, introduced by \cite{Anantharam87a} and recently studied by \cite{Komiyama15}.
%who essentially proved that existing algorithms can be easily extended to the multiple-play case, with provable guarantees on their regret.
% In the context of CR, this first extension can model a centralized agent taking decisions for end-devices, for instance a base station affecting devices to channels.
% Multiple-Play MAB algorithms successfully used for other Cognitive Radio tasks,
% for instance by \cite{Maghsudi16} for 5G small-cells or by \cite{Bonnefoi16} for Internet-of-Things (IoT) networks.
%
% Another generalization targets the opposite goal, where several devices
% are learning and conjointly accessing the same network,
% trying to find a consensus in a decentralized learning way.
% It is harder as there is a game theoretic-like situation, for instance where devices try to avoid collisions.
%
Due to the communication cost implied by a central controller, a more relevant model is the
\emph{decentralized multi-player} multi-armed bandit model, introduced by \cite{Zhao10} and \cite{Anandkumar10,Anandkumar11}, in which players select arms individually and collisions may occur, that yield a loss of reward.
Further algorithms were proposed in similar models by \cite{Tekin12IEEE} and \cite{Kalathil12} (under the assumption that each arm is a Markov chain)
and by \cite{Avner15,Avner16} and \cite{Rosenski16} (for \iid{} or piece-wise \iid{} arms).
%
The goal for every player is to select most of the time one of the $M$ best arms, without colliding too often with other players.
A first difficulty relies in the well-known trade-off between \emph{exploration} and \emph{exploitation}: players need to explore all arms to estimate their means while trying to focus on the best arms to gain as much rewards as possible.
The decentralized setting considers no exchange of information between players, that only know $K$ and $M$, and to avoid collisions, players should furthermore find orthogonal configurations (\ie, the $M$ players use the $M$ best arms without any collision), without communicating.
Hence, in that case the trade-off is to be found between exploration, exploitation \emph{and} low collisions.

@@

In this section, we take a step back and present the different feedback levels possible for multi-player MAB algorithms. For each of them, we propose algorithmic solutions supported by both experimental and theoretical guarantees. In the presence of sensing information, our contributions are a new problem-dependent regret lower bound, tighter than previous work, and the introduction of two algorithms, \RandTopM{} and \MCTopM{}. Both are shown to achieve an asymptotically optimal number of selections of the sub-optimal arms, and for \MCTopM{} we furthermore establish a logarithmic upper bound on the regret, that follows from a careful control of the number of collisions. In the absence of sensing information, we propose the \Selfish{} heuristic and investigate its performance. Our study of  this algorithm is supported by (promising) empirical performance and some first (disappointing) theoretical elements.

The rest of the section is organized as follows.
We introduce the multi-player bandit model with three feedback levels in Section~\ref{sec:5:model}, and give a new regret lower bound in Section~\ref{sec:5:lowerbound}.
The \RandTopM, \MCTopM{} and \Selfish{} algorithms are introduced in Section~\ref{sec:5:algorithms}, with the result of our experimental study reported in
Section~\ref{sec:5:experiments}. Theoretical elements are then presented in Section~\ref{sec:5:upperbounds}.
% and finally Section~\ref{sec:5:conclusion} concludes.

@@

% explain the interaction protocol and the GOAL
In the multi-player MAB setting, there are $M \in \{1,\dots,K\}$ players (or agents),
that have to make decisions at some pre-specified time instants.
At time step $t \in\mathbb{N},t\geq1$, player $j$ selects an arm $A^j(t)$, independently from the other players' selections.
A \emph{collision} occurs at time $t$ if at least two players choose the same arm.
We introduce the two events, for $j\in\{1,\dots,M\}$ and $k\in\{1,\dots,K\}$,
\begin{equation}
  C^j(t) :=  \{ \exists j' \neq j : A^{j'}(t) = A^j(t) \}
  \ \ \ \text{and} \ \ \ C_k(t) :=  \left\{ \# \{ j : A^j(t) = k\} > 1 \right\},
\end{equation}
that respectively indicate that a collision occurs at time $t$ for player $j$ and that a collision occurs at time $t$ on arm $k$.
Each player $j$ then receives (and observes) the \emph{binary rewards}
$r^j(t) \in \{0,1\}$,
\begin{equation}
  r^j(t) := Y_{A^j(t),t} \; \indic(\overline{C^j(t)}).
\end{equation}
In words, she receives the reward of the selected arm if she is the only one to select this arm, and a reward zero otherwise\footnote{This provides another reason to focus on the Bernoulli model. It is the hardest model, in the sense that receiving a reward zero is not enough to detect collisions. For other models, the data streams $(Y_{k,s})_s$ are usually continuously distributed, with no mass at zero. Hence receiving $r^j(t) = 0$ directly gives $\indic(C^j(t)) = 1$.}.
Other models for rewards loss have been proposed in the literature (\eg, the reward is randomly allocated to one of the players selecting it), but we focus on full reward occlusion in this Section.

A multi-player MAB strategy is a tuple $\rho = (\rho^1,\dots,\rho^M)$ of arm selection strategies for each player, and the goal is to propose a strategy that maximizes the total reward of the system, under some constraints.
First, each player $j$ should adopt a \emph{sequential} strategy $\rho^j$, that decides which arm to select at time $t$ based on \emph{previous observations}.
Previous observations for player $j$ at time $t$ always include the previously chosen arms $A^j(s)$ and received rewards $r^j(s)$ for $s<t$, but may also include the \emph{sensing information} $Y_{A^j(t),t}$ or the \emph{collision information} $C^j(t)$.
More precisely, depending on the application, one may consider the following three observation models, \modelun, \modeldeux{} and \modeltrois.
%
% If the arms have continuous distributions (or are such that  $\Pr(Y_{k,s}=0)=0$), the \emph{sensing information} $Y_{A^j(t),t}$ and \emph{collision information} $C^j(t)$ can always be extracted from the reward information. But for Bernoulli distribution, one may consider the following three observation models, that are not equivalent:
\begin{itemize}
  \item[\modelun]
    \textbf{Simultaneous sensing and collision}: player $j$ observes  $Y_{A^j(t),t}$ \emph{and} $C^j(t)$ (not previously studied).
  \item[\modeldeux]
    \textbf{Sensing, then collision}: player $j$ observes $Y_{A^j(t),t}$, \emph{then} observes the reward, and thus also $C^j(t)$ only if $Y_{A^j(t),t} = 1$.
    This common setup, studied for example by \cite{Anandkumar11,Avner15,Rosenski16}, is relevant to model the OSA problem: the device first checks for the presence of primary users in the chosen channel.
    If this channel is free ($Y_{A^j(t),t}=1$), the transmission is successful ($r^j(t)=1$) if no collision occurs with other smart devices ($\overline{C^j(t)}$).
  \item[\modeltrois]
    \textbf{No sensing}: player $j$ only observes the reward $r^j(t)$.
    For IoT networks, this reward can be interpreted as an acknowledgement from a Base Station,
    received when a communication was successful.
    A lack of acknowledgment may be due to a collision
    with a device from the background traffic $(Y_{A^j(t),t}=0)$,
    or to a collision with one of the others players ($C^j(t)$).
    However, the sensing and collision information are censored.
    Recently, our work presented in Chapter~\ref{chapter:4} \cite{Bonnefoi17} presented the first (bandit-based) algorithmic solutions under this (harder) feedback model, in a slightly different setup, more suited to large scale IoT applications. %\footnote{\cite{Bonnefoi17} considers a different setting, more suited for large-scale IoT networks, and a future work will be to extend our work to this setting with more than $K$ players but small probability of activations.}.
\end{itemize}
\noindent Under each of these three models, we define $\cF^j_t$ to be the filtration generated by the observations gathered by player $j$ up to time $t$ (which contains different information under models \modelun, \modeldeux{} and \modeltrois).
While a \emph{centralized} algorithm may select the vector of actions for all players $(A^1(t),\dots,A^M(t))$ based on all the observations from $\bigcup_j \cF^j_{t-1}$, under a \emph{decentralized} algorithm the arm selected at time $t$ by player $j$ only depends on the past observation of this player.
More formally, $A^j(t)$ is assumed to be $\cF^j_{t-1}$-measurable.

% [Index sets \Mworst, \Mbest]
\begin{definition}\label{def:5:MbestMworst}
  We denote by $\mu_1^*$ the best mean, $\mu_2^*$ the second best etc, and
  by \Mbest{} the (non-sorted) set of the indices of the $M$ arms with largest mean (\emph{best arms}): if $\mu_1^* = \mu_{k_1}, \dots, \mu_M^* = \mu_{k_M}$
  then $\Mbest = \{k_1, \dots, k_M\}$.
  %
  Similarly, \Mworst{} denotes the set of indices of the $K-M$ arms with smallest means (\emph{worst arms}),
  $\{1, \dots, K\} \setminus \Mbest$.
  %
  Note that they are both uniquely defined if $\mu_M^* > \mu_{M+1}^*$.
\end{definition}

Following a natural approach in the bandit literature, we evaluate the performance of a multi-player strategy using the \emph{expected regret} (later simply referred to as regret), that measures the performance gap with respect to the best possible strategy.
The regret of the strategy $\rho$ at horizon $T$ is the difference between the cumulated reward of an oracle strategy, assigning in this case the $M$ players to \Mbest,
and the cumulated reward of strategy $\rho$:
\begin{equation}\label{eq:5:regret}
  R_T(\boldsymbol{\mu}, M, \rho) := \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M r^j(t)\right].
\end{equation}
Maximizing the expected sum of global reward of the system is indeed equivalent to minimizing the regret, and we now investigate the best possible \emph{regret rate} of a decentralized multi-player algorithm.

@@

% \begin{definition}[]\label{def:5:nbCollisions}
  Let $\cC_k(T)$ be the \emph{number of colliding players}\footnote{When $n$ players choose arm $k$ at time $t$, this counts as $n$ collisions, not just one. So $\cC_k(T)$ counts the total \emph{number of colliding players} rather than the number of collision events. Hence there is small abuse of notation when calling it a number of collisions.}
  on arm $k\in\{1,\dots,K\}$ up to horizon $T$:
  \vspace*{-5pt}  % XXX
  \begin{equation}
    \cC_k(T) :=
    \sum_{t=1}^{T} \sum_{j=1}^{M} \indic(C^j(t)) \indic(A^j(t) = k).
  \end{equation}
\end{definition}

@@

\begin{lemma}\label{lem:5:DecompositionRegret}
  For any bandit instance $\boldsymbol{\mu}\in\cP_M$ such that $\mu_M^* > \mu_{M+1}^*$, it holds that
  \hfill{}
  (Proved in App.\ref{proof:5:DecompositionRegret})
  \vspace*{-5pt}  % XXX
  \begin{small}
    \begin{equation*}\label{eq:5:termeundeuxtermetrois}
      R_T(\boldsymbol{\mu}, M, \rho) =
      \underbrace{\sum_{k \in \Mworst} (\mu_M^* -  \mu_k) \E_{\mu}[T_k(T)]}_{\emph{\mytag{(a)}{eq:5:term1}}}
      + \underbrace{\sum_{k \in \Mbest} (\mu_k -  \mu_M^*) (T - \E_{\mu}[T_k(T)])}_{\emph{\mytag{(b)}{eq:5:term2}}}
      + \underbrace{\sum_{k=1}^{K} \mu_k \E_{\mu}[\cC_k(T)]}_{\emph{\mytag{(c)}{eq:5:term3}}}.
    \end{equation*}
  \end{small}
\end{lemma}

In this decomposition, term \ref{eq:5:term1} counts the lost rewards due to \emph{sub-optimal arms} selections ($k \in \Mworst$), term \ref{eq:5:term2} counts the number of times the \emph{best arms} were \emph{not} selected ($k \in \Mbest$), and term \ref{eq:5:term3} counts the weighted number of collisions, on \emph{all arms}. It is valid for both centralized and decentralized algorithms. For centralized algorithms, due to the absence of collisions, \ref{eq:5:term3} is obviously zero, and \ref{eq:5:term2} is non-negative, as $T_k(T) \leq T$. For decentralized algorithms, \ref{eq:5:term3} may be significantly large, and term \ref{eq:5:term2} may be negative, as many collisions on arm $k$ may lead to $T_k(T) > T$. However, a careful manipulation of this decomposition (see Appendix~\ref{proof:5:1stLowerBound}) shows that the regret is always lower bounded by term \ref{eq:5:term1}.

\begin{lemma}\label{lem:5:1stLowerBound}
    For any strategy $\rho$ and $\boldsymbol{\mu}\in\cP_M$, it holds that
    $ R_T(\boldsymbol{\mu}, M, \rho)    \geq \sum\limits_{k \in \Mworst} (\mu_M^*- \mu_k) \E_{\mu}[T_k(T)]$.
\end{lemma}

>>>>

% Multi-player Multi-Armed Bandits (MAB) have been extensively studied in the literature, motivated by applications to Cognitive Radio systems.
% Driven by such applications as well, we motivate the introduction of several levels of feedback for multi-player MAB algorithms.
% Existing works assume that \emph{sensing information} is available to the algorithm. Under this assumption, we improve the state-of-the-art lower bound for the regret of any decentralized algorithms (of a certain class).
% We introduce two algorithms, \RandTopM{} and \MCTopM{}, that are shown to empirically outperform existing algorithms. Moreover, we provide strong theoretical guarantees for these algorithms, including
% a notion of asymptotic optimality in terms of the number of selections of bad arms.
% We then introduce a promising heuristic, called \Selfish{}, that can operate
% without sensing information, which is crucial for emerging applications to Internet of Things networks. We investigate the empirical performance of this algorithm
% and provide some first theoretical elements for the understanding of its behavior.

@@

For the development of Cognitive Radio, a crucial step is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$).
With the presence of a central controller that can assign the devices to separate channels, this amounts to choosing at each time step \emph{several} arms of a MAB in order to maximize the global rewards, and can thus be viewed as an application of the multiple-play bandit, introduced by \cite{Anantharam87a} and recently studied by \cite{Komiyama15}.
They essentially proved that existing algorithms can be easily extended to the multiple-play case, with provable guarantees on their regret.
% In the context of CR, this first extension can model a centralized agent taking decisions for end-devices, for instance a base station affecting devices to channels.
% Multiple-Play MAB algorithms successfully used for other Cognitive Radio tasks,
% for instance by \cite{Maghsudi16} for 5G small-cells or by \cite{Bonnefoi16} for Internet-of-Things (IoT) networks.
%
% Another generalization targets the opposite goal, where several devices
% are learning and conjointly accessing the same network,
% trying to find a consensus in a decentralized learning way.
% It is harder as there is a game theoretic-like situation, for instance where devices try to avoid collisions.
%
Due to the communication cost implied by a central controller, a more relevant model is the
\emph{decentralized multi-player} multi-armed bandit model, introduced by \cite{Zhao10} and \cite{Anandkumar10,Anandkumar11}, in which players select arms individually and collisions may occur, that yield a loss of reward.
Further algorithms were proposed in similar models by \cite{Tekin12IEEE} and \cite{Kalathil12} (under the assumption that each arm is a Markov chain)
and by \cite{Avner15,Avner16} and \cite{Rosenski16} (for \iid{} or piece-wise \iid{} arms).
%
The goal for every player is to select most of the time one of the $M$ best arms, without colliding too often with other players.
A first difficulty relies in the well-known trade-off between \emph{exploration} and \emph{exploitation}: players need to explore all arms to estimate their means while trying to focus on the best arms to gain as much rewards as possible.
The decentralized setting considers no exchange of information between players, that only know $K$ and $M$, and to avoid collisions, players should furthermore find orthogonal configurations (\ie, the $M$ players use the $M$ best arms without any collision), without communicating.
Hence, in that case the trade-off is to be found between exploration, exploitation \emph{and} low collisions.

@@

In this chapter, we take a step back and present the different feedback levels possible for multi-player MAB algorithms. For each of them, we propose algorithmic solutions supported by both experimental and theoretical guarantees. In the presence of sensing information, our contributions are a new problem-dependent regret lower bound, tighter than previous work, and the introduction of two algorithms, \RandTopM{} and \MCTopM{}. Both are shown to achieve an asymptotically optimal number of selections of the sub-optimal arms, and for \MCTopM{} we furthermore establish a logarithmic upper bound on the regret, that follows from a careful control of the number of collisions. In the absence of sensing information, we propose the \Selfish{} heuristic and investigate its performance. Our study of this algorithm is supported by (promising) empirical performance and some first (disappointing) theoretical elements.

\paragraph{Outline.}
%
The rest of the chapter is organized as follows.
We introduce the multi-player bandit model with three feedback levels in Section~\ref{sec:5:model}, and give a new regret lower bound in Section~\ref{sec:5:lowerbound}.
The \RandTopM, \MCTopM{} and \Selfish{} algorithms are introduced in Section~\ref{sec:5:algorithms},
for which we present our theoretical analysis in Section~\ref{sec:5:upperbounds}.
Finally, the result of our experimental study reported in Section~\ref{sec:5:experiments}.
% and finally Section~\ref{sec:5:conclusion} concludes.

@@

% explain the interaction protocol and the GOAL
In the multi-player MAB setting, there are $M \in \{1,\dots,K\}$ players (or agents),
that have to make decisions at some pre-specified time instants.
At time step $t \in\mathbb{N},t\geq1$, player $j$ selects an arm $A^j(t)$, independently from the other players' selections.

\begin{definition}
  A \emph{collision} occurs at time $t$ if at least two players choose the same arm.
  We introduce the two events, for $j\in\{1,\dots,M\}$ and $k\in\{1,\dots,K\}$,
  \begin{equation}
    C^j(t) :=  \{ \exists j' \neq j : A^{j'}(t) = A^j(t) \}
    \ \ \ \text{and} \ \ \ C_k(t) :=  \left\{ \# \{ j : A^j(t) = k\} > 1 \right\},
  \end{equation}
  that respectively indicate that a collision occurs at time $t$ for player $j$ and that a collision occurs at time $t$ on arm $k$.
\end{definition}

Each player $j$ then receives (and observes) the \emph{binary rewards}
$r^j(t) \in \{0,1\}$,
\begin{equation}
  r^j(t) := Y_{A^j(t),t} \; \indic(\overline{C^j(t)}).
\end{equation}
In words, she receives the reward of the selected arm if she is the only one to select this arm, and a reward zero otherwise.
This provides another reason to focus on the Bernoulli model. It is the hardest model, in the sense that receiving a reward zero is not enough to detect collisions. For other models, the data streams $(Y_{k,s})_s$ are usually continuously distributed, with no mass at zero. Hence receiving $r^j(t) = 0$ directly gives $\indic(C^j(t)) = 1$.
Other models for rewards loss have been proposed in the literature, for instance the reward is randomly allocated to one of the players selecting it, but we focus on full reward occlusion in this chapter.

A multi-player MAB strategy is a tuple $\rho = (\rho^1,\dots,\rho^M)$ of arm selection strategies for of each of the $M$ players, and the goal is to propose a strategy that maximizes the total reward of the system, under some constraints.
First, each player $j$ should adopt a \emph{sequential} strategy $\rho^j$, that decides which arm to select at time $t$ based on \emph{previous observations}.
Previous observations for player $j$ at time $t$ always include the previously chosen arms $A^j(s)$ and received rewards $r^j(s)$ for $s<t$, but may also include the \emph{sensing information} $Y_{A^j(t),t}$ or the \emph{collision information} $C^j(t)$.
More precisely, depending on the application, one may consider the following three observation models, \modelun, \modeldeux{} and \modeltrois.

If the arms have continuous distributions (or are such that  $\Pr(Y_{k,s}=0)=0$), the \emph{sensing information} $Y_{A^j(t),t}$ and \emph{collision information} $C^j(t)$ can always be extracted from the reward information.
But for Bernoulli distributions, one may consider the following three observation models, that are not equivalent:

\begin{itemize}
  \item[\modelun]
    \textbf{Simultaneous sensing and collision}: player $j$ observes  $Y_{A^j(t),t}$ \emph{and} $C^j(t)$ (not previously studied).
  \item[\modeldeux]
    \textbf{Sensing, then collision}: player $j$ observes $Y_{A^j(t),t}$, \emph{then} observes the reward, and thus also $C^j(t)$ only if $Y_{A^j(t),t} = 1$.
    This common setup, studied for example by \cite{Anandkumar11,Avner15,Rosenski16}, is relevant to model the OSA problem: the device first checks for the presence of primary users in the chosen channel.
    If this channel is free ($Y_{A^j(t),t}=1$), the transmission is successful ($r^j(t)=1$) if no collision occurs with other smart devices ($\overline{C^j(t)}$).
  \item[\modeltrois]
    \textbf{No sensing}: player $j$ only observes the reward $r^j(t)$.
    For IoT networks, this reward can be interpreted as an acknowledgement from a Base Station,
    received when a communication was successful.
    A lack of acknowledgment may be due to a collision
    with a device from the background traffic $(Y_{A^j(t),t}=0)$,
    or to a collision with one of the others players ($C^j(t)$).
    However, the sensing and collision information are censored.
    Recently, our work presented in Chapter~\ref{chapter:4} \cite{Bonnefoi17} presented the first (bandit-based) algorithmic solutions under this (harder) feedback model, in a slightly different setup, more suited to large scale IoT applications. %\footnote{\cite{Bonnefoi17} considers a different setting, more suited for large-scale IoT networks, and a future work will be to extend our work to this setting with more than $K$ players but small probability of activations.}.
\end{itemize}

Under each of these three models, we define $\cF^j_t$ to be the filtration generated by the observations gathered by player $j$ up to time $t$ (which contains different information under models \modelun, \modeldeux{} and \modeltrois).
While a \emph{centralized} algorithm may select the vector of actions for all players $(A^1(t),\dots,A^M(t))$ based on all the observations from $\bigcup_j \cF^j_{t-1}$, under a \emph{decentralized} algorithm the arm selected at time $t$ by player $j$ only depends on the past observation of this player.
More formally, $A^j(t)$ is assumed to be $\cF^j_{t-1}$-measurable.

% [Index sets \Mworst, \Mbest]
\begin{definition}\label{def:5:MbestMworst}
  We denote by $\mu_1^*$ the best mean, $\mu_2^*$ the second best etc, and
  by \Mbest{} the (non-sorted) set of the indices of the $M$ arms with largest mean (\emph{best arms}): if $\mu_1^* = \mu_{k_1}, \dots, \mu_M^* = \mu_{k_M}$
  then $\Mbest = \{k_1, \dots, k_M\}$.
  %
  Similarly, \Mworst{} denotes the set of indices of the $K-M$ arms with smallest means (\emph{worst arms}),
  $\{1, \dots, K\} \setminus \Mbest$.

  Note that they are both uniquely defined if $\mu_M^* > \mu_{M+1}^*$.
\end{definition}

Following a natural approach in the bandit literature, we evaluate the performance of a multi-player strategy using the \emph{expected regret} (later simply referred to as regret), that measures the performance gap with respect to the best possible strategy.
The regret of the strategy $\rho$ at horizon $T$ is the difference between the cumulated reward of an oracle strategy, assigning in this case the $M$ players to \Mbest,
and the cumulated reward of strategy $\rho$:

\begin{definition}
  The excepted centralized multi-player regret is defined by
  \begin{equation}\label{eq:5:regret}
    R_T(\boldsymbol{\mu}, M, \rho) := \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M r^j(t)\right].
  \end{equation}
\end{definition}

With this definition, maximizing the expected sum of global reward of the system is indeed equivalent to minimizing the regret, and we investigate the best possible \emph{regret rate} of a decentralized multi-player algorithm in the next section.

@@

% \begin{definition}[]\label{def:5:nbCollisions}
  Let $\cC_k(T)$ be the \emph{number of colliding players} on arm $k\in\{1,\dots,K\}$ up to horizon $T$:
  % \vspace*{-5pt}  % XXX
  \begin{equation}
    \cC_k(T) :=
    \sum_{t=1}^{T} \sum_{j=1}^{M} \indic(C^j(t)) \indic(A^j(t) = k).
  \end{equation}
\end{definition}

Note that when $n$ players choose arm $k$ at time $t$, this counts as $n$ collisions, not just one. So $\cC_k(T)$ counts the total \emph{number of colliding players} rather than the number of collision events. Hence there is small abuse of notation when calling it a number of collisions.

@@

\begin{lemma}\label{lem:5:DecompositionRegret}
  For any bandit instance $\boldsymbol{\mu}\in\cP_M$ such that $\mu_M^* > \mu_{M+1}^*$, it holds that
  % \vspace*{-5pt}  % XXX
  % \begin{small}
    \begin{align}\label{eq:5:termeundeuxtermetrois}
      R_T(\boldsymbol{\mu}, M, \rho) &=
      \underbrace{\sum_{k \in \Mworst} (\mu_M^* -  \mu_k) \E_{\mu}[T_k(T)]}_{\emph{\mytag{(a)}{eq:5:term1}}} \\
      & \;\;\;\;\;\;
      + \;\; \underbrace{\sum_{k \in \Mbest} (\mu_k -  \mu_M^*) (T - \E_{\mu}[T_k(T)])}_{\emph{\mytag{(b)}{eq:5:term2}}}
      + \;\; \underbrace{\sum_{k=1}^{K} \mu_k \E_{\mu}[\cC_k(T)]}_{\emph{\mytag{(c)}{eq:5:term3}}}. \nonumber
    \end{align}
  % \end{small}
\end{lemma}

In this decomposition, term \ref{eq:5:term1} counts the lost rewards due to \emph{sub-optimal arms} selections ($k \in \Mworst$), term \ref{eq:5:term2} counts the number of times the \emph{best arms} were \emph{not} selected ($k \in \Mbest$), and term \ref{eq:5:term3} counts the weighted number of collisions, on \emph{all arms}.

\begin{proof}
  Using the definition of regret $R_T$ from \eqref{eq:5:regret}, and this collision indicator $\eta^j(t):=\indic(\overline{C^j(t)})$,
  \begin{align*}
    R(T)
    &= \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M Y_{A^j(t),t} \eta^j(t) \right]
     = \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M \mu_{A^j(t)} \eta^j(t)\right]
    \intertext{The last equality comes from the linearity of expectations, and the fact that $\E_{\mu}[Y_{k,t}] = \mu_k$ (for all $t$, from the \iid{} hypothesis), and the independence from $A^j(t)$, $\eta^j(t)$ and $Y_{k,t}$ (observed \emph{after} playing $A^j(t)$). So $\E_{\mu}[Y_{A^j(t),t} \eta^j(t)] = \sum_{k} \E_{\mu}[\mu_k \indic(A^j(t),t) \eta^j(t)] = \E_{\mu}[\mu_{A^j(t)} \eta^j(t)]$. And so}
    R(T)
    &= \E_{\mu}\left[ \sum_{t=1}^{T}\sum_{j \in \Mbest} \mu_j
      - \sum_{t=1}^{T} \sum_{j=1}^{M} \mu_{A^j(t)} \eta^j(t) \right] \\
      % &= \underbrace{\left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)}_{:= \overline{\mu}^*} (T M)
    &= \left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)
      - \sum_{k=1}^{K} \sum_{j=1}^{M} \mu_k \E_{\mu}\left[ T^j_k(T) \right]
      + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
    \intertext{For the first term, we have $T M = \sum\limits_{k=1}^{K} \sum\limits_{j=1}^{M} \E_{\mu}\left[ T^j_k(T)\right]$, and if we denote $\overline{\mu}^* := \frac{1}{M} \sum\limits_{j \in \Mbest} \mu_j$ the average mean of the $M$-best arms, then,}
    &= \sum_{k=1}^{K} \sum_{j=1}^{M} (\overline{\mu}^* - \mu_k) \E_{\mu}\left[ T^j_k(T) \right]
      + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
    \intertext{Let $\overline{\Delta_k} := \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $\{1,\dots,K\}$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
    &= \sum_{k \in \Mbest\setminus\{M\}} \overline{\Delta_k} \E_{\mu}\left[ T_k(T) \right]
      + \overline{\Delta_{M^*}} \E_{\mu}\left[ T_{M^*}(T) \right] \\
      &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} \overline{\Delta_k} \E_{\mu}\left[ T_k(T) \right]
      + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
    \intertext{But for $k = M^*$, $T_{M^*}(T) = T M^* - \sum\limits_{k \in \Mbest\setminus\{M\}} \E_{\mu}\left[ T_k(T) \right] - \sum\limits_{k \in \Mworst} \E_{\mu}\left[ T_k(T) \right]$, so by recombining the terms, we obtain,}
    &= \sum_{k \in \Mbest\setminus\{M\}} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ T_k(T) \right]
      + \overline{\Delta_{M^*}} T M^* \\
      &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ T_k(T) \right]
      + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
  \end{align*}
  The term $\overline{\Delta_k} - \overline{\Delta_{M^*}}$ simplifies to $\mu_{M^*} - \mu_k$, and so $\overline{\Delta_{M^*}} = \frac{1}{M} \sum_{k=1}^{M} \mu_k - \mu_{M^*}$ by definition of $\overline{\mu}^*$. And for $k=M^*$, $\mu_{M^*} - \mu_k = 0$, so the first sum can be written for $k = 1,\dots,M$ only, soIt
  % \vspace*{5pt}
  \begin{align*}
    R(T)
    &= \sum_{k \in \Mbest} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
      + \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) T \\
      &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
      + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right]
  \end{align*}
  And so we obtain the decomposition with three terms \ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3}.
  \begin{align*}
  R(T)
    & = \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) \left(T - \E_{\mu}\left[ T_k(T) \right]\right) \\
      & \;\; + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
      + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
  \end{align*}
  Which is exactly the decomposition we wanted to prove.
\end{proof}

The regret decomposition in Lemma~\ref{lem:5:DecompositionRegret} is valid for both centralized and decentralized algorithms.
For centralized algorithms, due to the absence of collisions, \ref{eq:5:term3} is obviously zero, and \ref{eq:5:term2} is non-negative, as $T_k(T) \leq T$. For decentralized algorithms, \ref{eq:5:term3} may be significantly large, and term \ref{eq:5:term2} may be negative, as many collisions on arm $k$ may lead to $T_k(T) > T$.
However, a careful manipulation of this decomposition shows that the regret is always lower bounded by term \ref{eq:5:term1}.

\begin{lemma}\label{lem:5:1stLowerBound}
    For any strategy $\rho$ and $\boldsymbol{\mu}\in\cP_M$, it holds that
    \begin{align*}
        R_T(\boldsymbol{\mu}, M, \rho)    \geq \sum\limits_{k \in \Mworst} (\mu_M^*- \mu_k) \E_{\mu}[T_k(T)].
    \end{align*}
\end{lemma}
